\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-06-30 paper.qmd
Add descriptions regarding five Gaussina cluster example

XXX THIS IS NOT INTERESTING IN ITS CURRENT FORM. THIS IS ONLY ANY INTERESTING EXAMPLE WHEN THE FILLED OUT VS FLAT SHAPES ARE DISCUSSED.

@@

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with tSNE in \pD{}, a video of the tour animation is available at (<https://youtu.be/RASEE7N5MbM>).
:::

@@

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with UMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/iG4bCPkJilw>).
:::

@@

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PaCMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/z07cKXi8EJQ>).
:::

In the \gD{} layouts generated by various NLDR techniques, as shown in @fig-NLDR-variety-gau, five well-separated clusters are shown. In tSNE (see @fig-NLDR-variety-gau (a)), these clusters appear closely. UMAP arranges all clusters in a parallel manner, with three aligned in one line and the other two in a separate line (see @fig-NLDR-variety-gau (b)). In contrast, PHATE shows two closely positioned clusters and three more distant ones (see @fig-NLDR-variety-gau (c)). In TriMAP, two clusters are close, though not as tightly as PHATE, while the other three are well-separated (see @fig-NLDR-variety-gau (d)). Finally, PaCMAP shows one central cluster and the remaining four spread out in different directions (see @fig-NLDR-variety-gau (e)).

<!-- five Gaussian clusters with different NLDR techniques-->
```{r}
#| warning: false
#| echo: false

## Import data
training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")

tsne_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_61.rds")
umap_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")
phate_data_gau <- read_rds("data/five_gau_clusters/phate_data_five_gau.rds")
trimap_data_gau <- read_rds("data/five_gau_clusters/trimap_data_five_gau.rds")
pacmap_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")

## Visualize embedding

nldr_gau1 <- tsne_data_gau |>
  ggplot(aes(x = tSNE1,
             y = tSNE2)) +
  geom_point(alpha=0.1, size=1) +
  interior_annotation("a")

nldr_gau2 <- umap_data_gau |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1) +
  interior_annotation("b")

nldr_gau3 <- phate_data_gau |>
  ggplot(aes(x = PHATE1,
             y = PHATE2)) +
  geom_point(alpha=0.1, size=1) +
  interior_annotation("c")

nldr_gau4 <- pacmap_data_gau |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2)) +
  geom_point(alpha=0.1, size=1) +
  interior_annotation("d")

nldr_gau5 <- trimap_data_gau |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2)) +
  geom_point(alpha=0.1, size=1) +
  interior_annotation("e")
```

```{r}
#| label: fig-NLDR-variety-gau
#| echo: false
#| fig-cap: "Five different NLDR representations of the same data. Different techniques and different parameter choices are used. Is there a best representation of the original data or are they all providing  equivalent information?"
#| fig-width: 8
#| fig-height: 2
#| out-width: 100%
#| fig-pos: H
# (a) tSNE (perplexity = 61), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2)
nldr_gau1 + nldr_gau2 + 
  nldr_gau3 + nldr_gau4 + nldr_gau5 + 
  plot_layout(ncol = 5)
```

```{r}
sc_xlims_gau <- c(-0.25, 1.25)
sc_ylims_gau <- c(-0.2, 1.2)
```

<!--trimesh with tsne-->
```{r}
gau1_scaled_obj <- gen_scaled_data(
  data = tsne_data_gau)
tsne_gau_scaled <- gau1_scaled_obj$scaled_nldr

## Compute hexbin parameters
num_bins_x_gau1 <- 12
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

gau1_model <- fit_highd_model(
  training_data = training_data_gau,
  emb_df = tsne_gau_scaled,
  bin1 = num_bins_x_gau1,
  r2 = r2_gau1,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "x"
)

df_bin_centroids_gau1 <- gau1_model$df_bin_centroids
df_bin_gau1 <- gau1_model$df_bin

## Triangulate bin centroids
tr1_object_gau1 <- tri_bin_centroids(
  df_bin_centroids_gau1, x = "c_x", y = "c_y")
tr_from_to_df_gau1 <- gen_edges(
  tri_object = tr1_object_gau1)

## Compute 2D distances
distance_gau1 <- cal_2d_dist(
  tr_coord_df = tr_from_to_df_gau1,
  start_x = "x_from",
  start_y = "y_from",
  end_x = "x_to",
  end_y = "y_to",
  select_vars = c("from", "to", "distance"))

## To find the benchmark value
benchmark_gau1 <- find_lg_benchmark(
  distance_edges = distance_gau1,
  distance_col = "distance")

trimesh_removed_gau1 <- vis_rmlg_mesh(
  distance_edges = distance_gau1, 
  benchmark_value = benchmark_gau1, 
  tr_coord_df = tr_from_to_df_gau1, 
  distance_col = "distance") +
  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
  interior_annotation("a", sc_ltr_pos) 
```

<!--trimesh with umap-->
```{r}
gau2_scaled_obj <- gen_scaled_data(
  data = umap_data_gau)
umap_gau_scaled <- gau2_scaled_obj$scaled_nldr

## Compute hexbin parameters
num_bins_x_gau2 <- 44
lim1 <- gau2_scaled_obj$lim1
lim2 <- gau2_scaled_obj$lim2
r2_gau2 <- diff(lim2)/diff(lim1)

gau2_model <- fit_highd_model(
  training_data = training_data_gau,
  emb_df = umap_gau_scaled,
  bin1 = num_bins_x_gau2,
  r2 = r2_gau2,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "x"
)

df_bin_centroids_gau2 <- gau2_model$df_bin_centroids
df_bin_gau2 <- gau2_model$df_bin

## Triangulate bin centroids
tr1_object_gau2 <- tri_bin_centroids(
  df_bin_centroids_gau2, x = "c_x", y = "c_y")
tr_from_to_df_gau2 <- gen_edges(
  tri_object = tr1_object_gau2)

## Compute 2D distances
distance_gau2 <- cal_2d_dist(
  tr_coord_df = tr_from_to_df_gau2,
  start_x = "x_from",
  start_y = "y_from",
  end_x = "x_to",
  end_y = "y_to",
  select_vars = c("from", "to", "distance"))

## To find the benchmark value
benchmark_gau2 <- find_lg_benchmark(
  distance_edges = distance_gau2,
  distance_col = "distance")

trimesh_removed_gau2 <- vis_rmlg_mesh(
  distance_edges = distance_gau2, 
  benchmark_value = benchmark_gau2, 
  tr_coord_df = tr_from_to_df_gau2, 
  distance_col = "distance") +
  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
  interior_annotation("b", sc_ltr_pos) 
```

<!--trimesh with phate-->
```{r}
# gau3_scaled_obj <- gen_scaled_data(
#   data = phate_data_gau)
# phate_gau_scaled <- gau3_scaled_obj$scaled_nldr
# 
# ## Compute hexbin parameters
# num_bins_x_gau3 <- 250
# lim1 <- gau3_scaled_obj$lim1
# lim2 <- gau3_scaled_obj$lim2
# r2_gau3 <- diff(lim2)/diff(lim1)
# 
# gau3_model <- fit_highd_model(
#   training_data = training_data_gau,
#   emb_df = phate_gau_scaled,
#   bin1 = num_bins_x_gau3,
#   r2 = r2_gau3,
#   is_bin_centroid = TRUE,
#   is_rm_lwd_hex = FALSE,
#   col_start_highd = "x"
# )
# 
# df_bin_centroids_gau3 <- gau3_model$df_bin_centroids
# df_bin_gau3 <- gau3_model$df_bin

df_bin_centroids_gau3 <- read_rds("data/five_gau_clusters/df_bin_centroids_gau_phate.rds")
df_bin_gau3 <- read_rds("data/five_gau_clusters/df_bin_gau_phate.rds")

## Triangulate bin centroids
tr1_object_gau3 <- tri_bin_centroids(
  df_bin_centroids_gau3, x = "c_x", y = "c_y")
tr_from_to_df_gau3 <- gen_edges(
  tri_object = tr1_object_gau3)

## Compute 2D distances
distance_gau3 <- cal_2d_dist(
  tr_coord_df = tr_from_to_df_gau3,
  start_x = "x_from",
  start_y = "y_from",
  end_x = "x_to",
  end_y = "y_to",
  select_vars = c("from", "to", "distance"))

## To find the benchmark value
benchmark_gau3 <- find_lg_benchmark(
  distance_edges = distance_gau3,
  distance_col = "distance")

trimesh_removed_gau3 <- vis_rmlg_mesh(
  distance_edges = distance_gau3, 
  benchmark_value = benchmark_gau3, 
  tr_coord_df = tr_from_to_df_gau3, 
  distance_col = "distance") +
  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
  interior_annotation("c", sc_ltr_pos) 
```

<!--trimesh with pacmap-->
```{r}
gau4_scaled_obj <- gen_scaled_data(
  data = pacmap_data_gau)
pacmap_gau_scaled <- gau4_scaled_obj$scaled_nldr

## Compute hexbin parameters
num_bins_x_gau4 <- 18
lim1 <- gau4_scaled_obj$lim1
lim2 <- gau4_scaled_obj$lim2
r2_gau4 <- diff(lim2)/diff(lim1)

gau4_model <- fit_highd_model(
  training_data = training_data_gau,
  emb_df = pacmap_gau_scaled,
  bin1 = num_bins_x_gau4,
  r2 = r2_gau4,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "x"
)

df_bin_centroids_gau4 <- gau4_model$df_bin_centroids
df_bin_gau4 <- gau4_model$df_bin

## Triangulate bin centroids
tr1_object_gau4 <- tri_bin_centroids(
  df_bin_centroids_gau4, x = "c_x", y = "c_y")
tr_from_to_df_gau4 <- gen_edges(
  tri_object = tr1_object_gau4)

## Compute 2D distances
distance_gau4 <- cal_2d_dist(
  tr_coord_df = tr_from_to_df_gau4,
  start_x = "x_from",
  start_y = "y_from",
  end_x = "x_to",
  end_y = "y_to",
  select_vars = c("from", "to", "distance"))

## To find the benchmark value
benchmark_gau4 <- find_lg_benchmark(
  distance_edges = distance_gau4,
  distance_col = "distance")

trimesh_removed_gau4 <- vis_rmlg_mesh(
  distance_edges = distance_gau4, 
  benchmark_value = benchmark_gau4, 
  tr_coord_df = tr_from_to_df_gau4, 
  distance_col = "distance") +
  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
  interior_annotation("d", sc_ltr_pos) 
```

<!--trimesh with trimap-->
```{r}
gau5_scaled_obj <- gen_scaled_data(
  data = trimap_data_gau)
trimap_gau_scaled <- gau5_scaled_obj$scaled_nldr

## Compute hexbin parameters
num_bins_x_gau5 <- 25
lim1 <- gau5_scaled_obj$lim1
lim2 <- gau5_scaled_obj$lim2
r2_gau5 <- diff(lim2)/diff(lim1)

gau5_model <- fit_highd_model(
  training_data = training_data_gau,
  emb_df = trimap_gau_scaled,
  bin1 = num_bins_x_gau5,
  r2 = r2_gau5,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "x"
)

df_bin_centroids_gau5 <- gau5_model$df_bin_centroids
df_bin_gau5 <- gau5_model$df_bin

## Triangulate bin centroids
tr1_object_gau5 <- tri_bin_centroids(
  df_bin_centroids_gau5, x = "c_x", y = "c_y")
tr_from_to_df_gau5 <- gen_edges(
  tri_object = tr1_object_gau5)

## Compute 2D distances
distance_gau5 <- cal_2d_dist(
  tr_coord_df = tr_from_to_df_gau5,
  start_x = "x_from",
  start_y = "y_from",
  end_x = "x_to",
  end_y = "y_to",
  select_vars = c("from", "to", "distance"))

## To find the benchmark value
benchmark_gau5 <- find_lg_benchmark(
  distance_edges = distance_gau5,
  distance_col = "distance")
benchmark_gau5 <- 0.2

trimesh_removed_gau5 <- vis_rmlg_mesh(
  distance_edges = distance_gau5, 
  benchmark_value = benchmark_gau5, 
  tr_coord_df = tr_from_to_df_gau5, 
  distance_col = "distance") +
  xlim(sc_xlims_gau) + ylim(sc_ylims_gau) +
  interior_annotation("e", sc_ltr_pos) 
```

<!--add trimesh for all five-->

```{r}
#| label: fig-trimesh-gau
#| echo: false
#| fig-cap: "Model generated with five different NLDR methods in $2\\text{-}D$ with approximately $65$ non-empty bins in each."
#| fig-width: 8
#| fig-height: 2
#| out-width: 100%
#| fig-pos: H

trimesh_removed_gau1 + trimesh_removed_gau2 +
  trimesh_removed_gau3 + trimesh_removed_gau4 + trimesh_removed_gau5 +
  plot_layout(ncol = 5)
```

<!--add three screenshots from each technique with you tube links-->

To investigate which is the reasonable representation to visualize the five spherical Gaussian cluster data or all NLDR methods provide equivalent information, we visualize all the models in \pD{} space. Models from all NLDR methods show five well-separated clusters (see @fig-gau1-sc, @fig-gau2-sc, @fig-gau3-sc, @fig-gau4-sc, and @fig-gau5-sc). This suggests that for the five Gaussian cluster dataset, all NLDR methods effectively preserve the global structure. tSNE displays clusters with varying densities, indicating their ability to capture within-cluster variation (see @fig-gau1-sc). On the other hand, both UMAP, PHATE, PaCMAP and TriMAP show clusters with flat surfaces, suggesting a failure to capture within-cluster variation (see @fig-gau2-sc, @fig-gau3-sc, @fig-gau4-sc and @fig-gau5-sc). Therefore, UMAP, PHATE, PaCMAP and TriMAP do not capture the local structure as effectively as other methods.

::: {#fig-gau1-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/sc_tsne_1.png)

![](figures/five_gau_clusters/sc_tsne_2.png)

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with tSNE in \pD{}, a video of the tour animation is available at (<https://youtu.be/RASEE7N5MbM>).
:::

::: {#fig-gau2-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/sc_umap_1.png)

![](figures/five_gau_clusters/sc_umap_2.png)

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with UMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/iG4bCPkJilw>).
:::

::: {#fig-gau3-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/sc_phate_1.png)

![](figures/five_gau_clusters/sc_phate_2.png)

![](figures/five_gau_clusters/sc_phate_3.png)

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PHATE in \pD{}, a video of the tour animation is available at (<https://youtu.be/L_PVLGwfOS0>).
:::

::: {#fig-gau4-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/sc_pacmap_1.png)

![](figures/five_gau_clusters/sc_pacmap_3.png)

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with PaCMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/z07cKXi8EJQ>).
:::

::: {#fig-gau5-sc layout-ncol="3" fig-pos="H"}
![](figures/five_gau_clusters/sc_trimap_1.png)

![](figures/five_gau_clusters/sc_trimap_2.png)

![](figures/five_gau_clusters/sc_trimap_3.png)

Screen shots of the **langevitour** of the five Gaussian clusters dataset, shows the model with TriMAP in \pD{}, a video of the tour animation is available at (<https://youtu.be/Chs1lYAoX2w>).
:::

When compare the NLDR representations and generated models, tSNE with perplexity $61$ appears to be a reasonable representation for visualizing the five Gaussian cluster dataset. This is supported by investigating the model generated with tSNE in the data space, which provides evidence that it preserves both local and global structures. Also, the NLDR representation with tSNE shows five well-separated clusters.

>>>>

<!-- XXX THIS IS NOT INTERESTING IN ITS CURRENT FORM. THIS IS ONLY ANY INTERESTING EXAMPLE WHEN THE FILLED OUT VS FLAT SHAPES ARE DISCUSSED. -->

The \gD{} layouts generated by tSNE, UMAP, and PaCMAP show five well-separated clusters which evident these methods effectively preserve the global structure. In tSNE (@fig-gau-tsne-sc (a)), these clusters appear closely. UMAP arranges all clusters in a parallel manner, with three aligned in one line and the other two in a separate line (@fig-gau-umap-sc (a)). In contrast, PaCMAP shows one central cluster and the remaining four spread out in different directions (@fig-gau-pacmap-sc (a)).

The tSNE and UMAP shows *filled out* clusters which provide evidence that these methods preserve the local structure (@fig-gau-tsne-sc (c) and @fig-gau-umap-sc (c)). On the other hand, PaCMAP shows *flat* shapes clusters in the model and evident that PaCMAP fail to capture the within-cluster variation (@fig-gau-pacmap-sc (c)).

@@

The tSNE layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data. The model fits the separation and tries to *filled out* the clusters. (The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/RASEE7N5MbM>). 
:::

@@

The UMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data. The model fits the separation and tries to *filled out* the clusters, but not as much as tSNE. (The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/iG4bCPkJilw>). 
:::

@@

The PaCMAP layout, model in \gD{}, and a view of the fit in projections from $4\text{-}D$, for the five Gaussian cluster data. The model fits the separation and shows *flat* shaped clusters. (The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/z07cKXi8EJQ>). 
:::

### 2024-06-30 paper.tex
Add descriptions regarding five Gaussina cluster example

XXX THIS IS NOT INTERESTING IN ITS CURRENT FORM. THIS IS ONLY ANY
INTERESTING EXAMPLE WHEN THE FILLED OUT VS FLAT SHAPES ARE DISCUSSED.

@@

\caption{\label{fig-gau-tsne-sc}Screen shots of the \textbf{langevitour}
of the five Gaussian clusters dataset, shows the model with tSNE in
\pD{}, a video of the tour animation is available at
(\url{https://youtu.be/RASEE7N5MbM}).}

@@

\caption{\label{fig-gau-umap-sc}Screen shots of the \textbf{langevitour}
of the five Gaussian clusters dataset, shows the model with UMAP in
\pD{}, a video of the tour animation is available at
(\url{https://youtu.be/iG4bCPkJilw}).}

@@

\caption{\label{fig-gau-pacmap-sc}Screen shots of the
\textbf{langevitour} of the five Gaussian clusters dataset, shows the
model with PaCMAP in \pD{}, a video of the tour animation is available
at (\url{https://youtu.be/z07cKXi8EJQ}).}

@@

In the \gD{} layouts generated by various NLDR techniques, as shown in
Figure~\ref{fig-NLDR-variety-gau}, five well-separated clusters are
shown. In tSNE (see Figure~\ref{fig-NLDR-variety-gau} (a)), these
clusters appear closely. UMAP arranges all clusters in a parallel
manner, with three aligned in one line and the other two in a separate
line (see Figure~\ref{fig-NLDR-variety-gau} (b)). In contrast, PHATE
shows two closely positioned clusters and three more distant ones (see
Figure~\ref{fig-NLDR-variety-gau} (c)). In TriMAP, two clusters are
close, though not as tightly as PHATE, while the other three are
well-separated (see Figure~\ref{fig-NLDR-variety-gau} (d)). Finally,
PaCMAP shows one central cluster and the remaining four spread out in
different directions (see Figure~\ref{fig-NLDR-variety-gau} (e)).

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-variety-gau-1.pdf}

\caption{\label{fig-NLDR-variety-gau}Five different NLDR representations
of the same data. Different techniques and different parameter choices
are used. Is there a best representation of the original data or are
they all providing equivalent information?}

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-trimesh-gau-1.pdf}

\caption{\label{fig-trimesh-gau}Model generated with five different NLDR
methods in \(2\text{-}D\) with approximately \(65\) non-empty bins in
each.}

To investigate which is the reasonable representation to visualize the
five spherical Gaussian cluster data or all NLDR methods provide
equivalent information, we visualize all the models in \pD{} space.
Models from all NLDR methods show five well-separated clusters (see
Figure~\ref{fig-gau1-sc}, Figure~\ref{fig-gau2-sc},
Figure~\ref{fig-gau3-sc}, Figure~\ref{fig-gau4-sc}, and
Figure~\ref{fig-gau5-sc}). This suggests that for the five Gaussian
cluster dataset, all NLDR methods effectively preserve the global
structure. tSNE displays clusters with varying densities, indicating
their ability to capture within-cluster variation (see
Figure~\ref{fig-gau1-sc}). On the other hand, both UMAP, PHATE, PaCMAP
and TriMAP show clusters with flat surfaces, suggesting a failure to
capture within-cluster variation (see Figure~\ref{fig-gau2-sc},
Figure~\ref{fig-gau3-sc}, Figure~\ref{fig-gau4-sc} and
Figure~\ref{fig-gau5-sc}). Therefore, UMAP, PHATE, PaCMAP and TriMAP do
not capture the local structure as effectively as other methods.

\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_tsne_1.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_tsne_2.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_tsne_3.png}\end{minipage}%

\caption{\label{fig-gau1-sc}Screen shots of the \textbf{langevitour} of
the five Gaussian clusters dataset, shows the model with tSNE in \pD{},
a video of the tour animation is available at
(\url{https://youtu.be/RASEE7N5MbM}).}

\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_umap_1.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_umap_2.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_umap_3.png}\end{minipage}%

\caption{\label{fig-gau2-sc}Screen shots of the \textbf{langevitour} of
the five Gaussian clusters dataset, shows the model with UMAP in \pD{},
a video of the tour animation is available at
(\url{https://youtu.be/iG4bCPkJilw}).}

\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_phate_1.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_phate_2.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_phate_3.png}\end{minipage}%

\caption{\label{fig-gau3-sc}Screen shots of the \textbf{langevitour} of
the five Gaussian clusters dataset, shows the model with PHATE in \pD{},
a video of the tour animation is available at
(\url{https://youtu.be/L_PVLGwfOS0}).}

\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_pacmap_1.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_pacmap_2.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_pacmap_3.png}\end{minipage}%

\caption{\label{fig-gau4-sc}Screen shots of the \textbf{langevitour} of
the five Gaussian clusters dataset, shows the model with PaCMAP in
\pD{}, a video of the tour animation is available at
(\url{https://youtu.be/z07cKXi8EJQ}).}

\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_trimap_1.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_trimap_2.png}\end{minipage}%
%
\begin{minipage}{0.33\linewidth}
\includegraphics{figures/five_gau_clusters/sc_trimap_3.png}\end{minipage}%

\caption{\label{fig-gau5-sc}Screen shots of the \textbf{langevitour} of
the five Gaussian clusters dataset, shows the model with TriMAP in
\pD{}, a video of the tour animation is available at
(\url{https://youtu.be/Chs1lYAoX2w}).}

When compare the NLDR representations and generated models, tSNE with
perplexity \(61\) appears to be a reasonable representation for
visualizing the five Gaussian cluster dataset. This is supported by
investigating the model generated with tSNE in the data space, which
provides evidence that it preserves both local and global structures.
Also, the NLDR representation with tSNE shows five well-separated
clusters.

>>>>

The \gD{} layouts generated by tSNE, UMAP, and PaCMAP show five
well-separated clusters which evident these methods effectively preserve
the global structure. In tSNE (Figure~\ref{fig-gau-tsne-sc} (a)), these
clusters appear closely. UMAP arranges all clusters in a parallel
manner, with three aligned in one line and the other two in a separate
line (Figure~\ref{fig-gau-umap-sc} (a)). In contrast, PaCMAP shows one
central cluster and the remaining four spread out in different
directions (Figure~\ref{fig-gau-pacmap-sc} (a)).

The tSNE and UMAP shows \emph{filled out} clusters which provide
evidence that these methods preserve the local structure
(Figure~\ref{fig-gau-tsne-sc} (c) and Figure~\ref{fig-gau-umap-sc} (c)).
On the other hand, PaCMAP shows \emph{flat} shapes clusters in the model
and evident that PaCMAP fail to capture the within-cluster variation
(Figure~\ref{fig-gau-pacmap-sc} (c)).

@@

\caption{\label{fig-gau-tsne-sc}The tSNE layout, model in \gD{}, and a
view of the fit in projections from \(4\text{-}D\), for the five
Gaussian cluster data. The model fits the separation and tries to
\emph{filled out} the clusters. (The \textbf{langevitour} software is
used to view the data with a tour, and the full video is available at
(\url{https://youtu.be/RASEE7N5MbM}).}

@@

\caption{\label{fig-gau-umap-sc}The UMAP layout, model in \gD{}, and a
view of the fit in projections from \(4\text{-}D\), for the five
Gaussian cluster data. The model fits the separation and tries to
\emph{filled out} the clusters, but not as much as tSNE. (The
\textbf{langevitour} software is used to view the data with a tour, and
the full video is available at (\url{https://youtu.be/iG4bCPkJilw}).}

@@

\caption{\label{fig-gau-pacmap-sc}The PaCMAP layout, model in \gD{}, and
a view of the fit in projections from \(4\text{-}D\), for the five
Gaussian cluster data. The model fits the separation and shows
\emph{flat} shaped clusters. (The \textbf{langevitour} software is used
to view the data with a tour, and the full video is available at
(\url{https://youtu.be/z07cKXi8EJQ}).}

@@
